---
title: "p8105_hw2_xl3712"
author: "Xieyu Liu"
date: "2026-10-02"
output: github_document
---

```{r}
library(tidyverse)
```

## Problem 1
```{r}
Transit_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")|>
  janitor::clean_names()|>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(entry = case_match(
    entry,
    "YES" ~ T,
    "NO" ~ F
  ))
```

The dataset contains line, station name, station latitude, station longitude, routes 1 to 11, entry, vending, entrance type and ADA compliant. I first imported the NYC Transit Data, and cleaned the variable names. Then, I select the relevant columns and converted the entry variable from "YES" or "NO" to TRUE or FALSE. The dataset has `r nrow(Transit_df)` rows and `r ncol(Transit_df)` columns. These data are not tidy because the routes served are stored across 11 columns instead of one variable.

```{r}
Transit_df |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct stations.

```{r}
Transit_df |>
  distinct(line, station_name, ada) |>
  filter(ada == "TRUE") |>
  nrow()
```

There are 84 ADA-compliant stations.

```{r}
Transit_df |>
  filter(vending == "NO")|>
  summarize(
    proportion = mean(entry)
  )
```

Among station entrance/exits without vending, 37.7% allow entrance.

```{r}
Transit_df_reformat = Transit_df |>
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

```{r}
Transit_df_reformat|>
  filter(route_name == "A")|>
  distinct(line, station_name) |>
  nrow()
```

There are 60 distinct stations that serve the A train.

```{r}
Transit_df_reformat|>
  filter(route_name == "A")|>
  distinct(line, station_name, ada) |>
  filter(ada == "TRUE") |>
  nrow()
```

Of these stations, 17 are ADA compliant.

## Problem 2

```{r}
library(readxl)
```


```{r}
Mr_Trash_Wheel_df = 
  read_excel("data/202509 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             skip = 1,
             range = cell_cols(1:14)
             )|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(
    sports_balls = as.integer(round(sports_balls))
    )
```

```{r}
Professor_Trash_Wheel_df = 
  read_excel("data/202509 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             skip = 1
             ) |>
  janitor::clean_names()|>
  filter(!is.na(dumpster))
```

```{r}
Gwynnda_df = 
  read_excel("data/202509 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynns Falls Trash Wheel",
             skip = 1
             )|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))
```

```{r}
Mr_Trash_Wheel_df = 
  Mr_Trash_Wheel_df|>
  mutate(trash_wheel = "Mr",
         year = as.character(year))

Professor_Trash_Wheel_df =
  Professor_Trash_Wheel_df|>
  mutate(trash_wheel = "Professor",
         year = as.character(year))

Gwynnda_df = 
  Gwynnda_df|>
  mutate(trash_wheel = "Gwynnda",
         year = as.character(year))

Trash_Wheel_df = 
  bind_rows(
    Mr_Trash_Wheel_df,
    Professor_Trash_Wheel_df,
    Gwynnda_df
            )|>
  relocate(trash_wheel)
```

The resulting dataset contains `r nrow(Trash_Wheel_df)` observations and includes information on trash wheel type, dumpster number, collection date, and the amount and type of trash collected. Key variables include `trash_wheel`, `dumpster`, `date`, `weight_tons`, `volume_cubic_yards`, and variables measuring different types, such as `plastic_bottles` and `polystyrene`.

```{r}
Trash_Wheel_df|>
  filter(trash_wheel == "Professor")|>
  summarise(total_weight = sum(weight_tons))
```

The total weight of trash collected by Professor Trash Wheel was 282 tons.

```{r}
Trash_Wheel_df|>
  filter(trash_wheel == "Gwynnda",
         month == "June",
         year == "2022")|>
  summarise(total_number_of_cigarette_butts = sum(cigarette_butts))
```

The total number of cigarette butts collected by Gwynnda in June 2022 was 18120.